---
title: "Logistic Regression"
description: |
  Logistic Regression on 2 species of Floridian Palmettos.
author:
  - name: Shelby Walsh
date: 03-13-2021
theme: style.css
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(tidyr)
library(lubridate)
library(tsibble)
library(feasts)
library(GGally)
library(broom)
library(jtools)
library(scales)
library(kableExtra)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
```


```{r, include = FALSE}
#read in data
palmetto <- read_csv("palmetto.csv", 
                     col_types = cols(.default = 'c')) %>%
  mutate(full_name = case_when(
    species == '1' ~ 'Serenoa repens',
    species == '2' ~ 'Sabal etonia'
  )) %>% 
  mutate(height = as.numeric(height)) %>% 
  mutate(length = as.numeric(length)) %>% 
  mutate(width = as.numeric(width)) %>% 
  mutate(green_lvs = as.numeric(green_lvs)) 


str(palmetto)

palmetto <- palmetto %>% 
  select(species, height, length, width, green_lvs, full_name)

str(palmetto$species)

palmetto$species <- as.factor(palmetto$species)

```

## Data Visualizations to explore Palmetto Species differences
```{r, fig.cap= "**Figure 1.** Exploritory visualization to explore the differences in two dominant Palmetto species, Sabel etonia and Serenoa repens, based on height and canopy length."}
palmetto_height <- ggplot(data = palmetto, aes(x = height, y = length)) +
                            geom_point(aes(color = full_name)) +
  scale_colour_manual(values = c("darkolivegreen4", "goldenrod2")) +
  labs(x = "Height (cm)",
       y = "Length (cm)", 
       title = "Comparison of Canopy Height and Length of Two South-Central \nFlorida Palmetto Species, Sabel etonia and Serenoa repens") +
  theme(legend.title = element_blank()) +
  theme(plot.title = element_text(size = 12,
                                   face = "bold",
        hjust = 0.5))

palmetto_height
```
- Height and length have a correlation with species

```{r, fig.cap= "**Figure 2.** Exploritory visualization to explore the differences in two dominant Palmetto species, Sabel etonia and Serenoa repens, based on canopy width and green leaf observations."}
palmetto_canopy <- ggplot(data = palmetto, aes(x = width, y = green_lvs)) +
                            geom_point(aes(color = full_name))+
  scale_colour_manual(values = c("darkolivegreen4", "goldenrod2"))+
  labs(x = "Width (cm)",
       y = "Green Leaf Count", 
       title = "Comparison of Canopy Width and Green Leaf Count of Two South-Central \nFlorida Palmetto Species, Sabel etonia and Serenoa repens") +
  theme(legend.title = element_blank()) +
  theme(plot.title = element_text(size = 12,
                                   face = "bold",
        hjust = 0.5))


palmetto_canopy
```
- There seems to be a difference between the 2 species based on green leaf count and width

```{r, fig.cap= "**Figure 3.** Exploritory visualization to explore the two strongest differences in two dominant Palmetto species, Sabel etonia and Serenoa repens, based on height and green leaf observations."}
palmetto_hg <- ggplot(data = palmetto, aes(x = green_lvs, y = height)) +
                            geom_point(aes(color = full_name)) +
  scale_colour_manual(values = c("darkolivegreen4", "goldenrod2"))+
  labs(x = "Green Leaf Count",
       y = "Height (cm)", 
       title = "Comparison of Height and Green Leaf Count of Two South-Central \nFlorida Palmetto Species, Sabel etonia and Serenoa repens") +
  theme(legend.title = element_blank()) +
  theme(plot.title = element_text(size = 12,
                                   face = "bold",
        hjust = 0.5))


palmetto_hg
```
- Green leaf count seem to be strong indicator for species

## Binary Logistic Regression for Palmetto Species, Sabel etonia and Serenoa repens. 
#### Specifically looking at the relationship between species identification and canopy height, length, width and green leaf observations.
```{r}
#running the regression and making a table
palmetto_blr <- glm(species ~ height + length + width + green_lvs,
                            data = palmetto,
                            family = "binomial")

#summary(palmetto_blr)

blr_tidy <- broom::tidy(palmetto_blr)

blr_tidy %>%
kbl(caption = "**Table 1.** Palmetto Logistic Regression Results") %>%
  kable_classic(full_width = F, html_font = "Cambria") %>% 
  add_header_above(c( "Species" = 5))



#ggplot(data = palmetto, aes(x = species, y = height)) +
  #geom_jitter(aes(color = length)) +
  #theme(legend.position = "none")
```

```{r}
#creating a clean table and labels
#clean_table <- tab_model(palmetto_blr)

#clean_table
```


## Model Prediction and Classification
#### Analyzing our regression model and looking at the accuracy of the predictions. Predictions are classified as correct when the probability is >0.5 of matching the orginal species.
```{r}
#predict species
blr_fitted <- palmetto_blr %>% 
 broom::augment(type.predict = "response") %>% 
  mutate(predict_species = ifelse(.fitted >= 0.5, "Serenoa repens", "Sabal etonia"))

#sort and predict species to find how often it is right
blr_correct <- blr_fitted %>% 
  mutate(species = case_when(species == 1 ~ "Serenoa repens",
                             species == 2 ~ "Sabal etonia")) %>% 
  mutate(correct = case_when(predict_species == species ~ "incorrect",
                             TRUE ~ "correct")) %>% 
  select(species, .fitted, predict_species, correct) %>% 
 group_by(species, correct) %>% 
  summarize(number_correct = n())

#pivot to get table in tidy format
blr_correct <- blr_correct %>% 
  pivot_wider(names_from = correct,
              values_from = number_correct)

pretty_correct <- blr_correct %>% 
  group_by(species) %>% 
mutate(percent_correct = correct/(correct + incorrect)) %>% 
  mutate(percent_correct = percent(percent_correct))

pretty_correct <- pretty_correct %>% 
  rename("Species" = species,
         "Correct" = correct,
         "Incorrect" = incorrect, 
         "% Correct" = percent_correct)

pretty_correct %>%
kbl(caption = "**Table 2.** Predicted Palmetto Species Identification Results") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

